What Can Stack Traces Tell us about the Use of Exceptions in Java?

### 2014-05-11 android-stacks.tex
monads more ideas - reduced
Proof reading comments

Modern applications have to cope with an increasing number of abnormal
computation states that arise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware. The exception handling
mechanism~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions. Although the
exception handling mechanism have been embedded in several programming languages
(e.g. Java, C++, C\#) and the target of several studies
(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
the exception handling code is often generally poorly understood and the least
tested part of software systems.

Often some exception handling constructions may lead the developers into believing that
by just re-throwing the exceptions they can forget about the exceptional
situations during the development of the "happy path". This "ignore-for-now"
approach may turn the exception handling into a generalized "goto"
mechanism~\cite{mandrioli1992advances} making the program more complex and even
less reliable. This behaviour may lead to the well known uncaught exceptions
problem~\cite{jo2004uncaught}, one of the main causes of application crashes.

Moreover, some languages use exceptions to represent programming mistakes
(out-of-bounds array index, division-by-zero, access to a null). In these cases
exceptions are implicitly signaled by the runtime environment when such
condition happens. One of such implicit exceptions (caused by data conversion
from a 64-bit floating point to a 16-bit signed integer) was the responsible for
the famous failure on Ariane 5's first test flight~\cite{lions1996ariane} ---
leading to the rocket self-destructing 37 seconds after launch and a loss of 500
million dollars. Besides Ariane 5, several applications crash everyday due to
uncaught exceptions~\cite{jo2004uncaught}. The consequences of not careful
exception usage may have the opposite effect of its initial intention of
improving system robustness.

@@

To answer these questions, we conduct an exploratory study of stack traces in
the wild. Using a custom tool called ExceptionMiner, that was specifically
developed to for this study, we mined stack traces from issues across all Java
projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
from which 28,800 stack traces were extracted and used as a source of
information for our analysis. The stack trace analysis  was augmented by
additional information extracted using byte code and source code analyses done
on a carefully selected subset of our initial sample (482 Android projects).
%This study was divided in two phases: in the fist phase all the issues of all
%Java projects available on GitHub were analyzed and the embedded stack traces
%extracted using ExceptionMiner.

@@

  \item  It briefly introduces ExceptionMiner, a tool developed to support the
    analysis.

@@

The contributions of this work allow for developers of robust Java applications
to familiarize with the most common reported exceptions mis-uses which is the
first step to help developers to avoid making them; (ii) designers of languages
to consider ways of reducing the abundance of NullPointerExceptions, (iii) and
tool designers to consider developing tools that enable a cross-project defect
analysis. We strongly believe that, specially in the open-source environment,
faults are not to be hidden in a private bug issue. Faults should be share and
discussed, so that a developer can learn from other projects mistakes.
Currently, the search facilities of repositories are very limited, and the
ExceptionMiner tool is a contribution in this direction.

@@

The Java specification~\cite{gosling2000java} suggests that user-defined
exceptions should be checked exceptions. The main reason is because doing so the
callers of a method will know about the exceptions that a method can throw, and
so they can decide what to do about them. However, there is a long-lasting
debate about the use of checked and unchecked exceptions in
Java~\cite{javatut,stackoverlow,debate}\footnote{152 questions in Stackoverlow
are related to this debate}, since there are pros and cons associated to each of
them.

@@

Several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. 
A compiled list of the guidelines related to exception types and exception
propagation in Java the following: \footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

@@

\textbf{I-Use checked exceptions to represent recoverable
conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
The developer should use checked exceptions for conditions from which the caller
is expected to recover. By confronting the API user with a checked exception,
the API designer is telling the client to handle the exceptional condition. The
client can explicitly ignore the exception (swallowing, or converting it to
other type) at the expense of the program's robustness~\cite{gosling2000java}.

@@

%\noindent\emph{Exception Documentation}
\textbf{IV-Document all exceptions thrown by a
method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly signaled runtime exceptions\footnote{This
excludes the implicitly signaled JVM runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of the exceptions the method can throw. If the developer fails to
do follow this practice (specially when developing library code) it will be
difficult or even impossible for the caller to make effective use of such method
~\cite{wirfs2006toward, bloch2008effective}.

@@

This work describes an exploratory study which was based on a sequential
mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and
integrating both quantitative and qualitative analysis along the stages of the
research process. The main goal of this study was to gain a better understanding
of the use of exceptions in Java programs based mainly on the information
available on exception stack traces. More specifically, we aimed at answering
the following research questions:

@@

For RQ1 and RQ2, we explore the domain quantitatively, and use theirs answers as
a basis to investigate the third research question. RQ3 aimed at finding whether
of exception (mis)-uses could be revealed through stack trace analysis. To
support this investigation, source code and bytecode analysis was used to
leverage the  understanding of stack traces. For RQ3 we investigate
quantitatively and highlight interesting cases by exploring cases qualitatively.
Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
conducted to answer these questions.

@@

We begun our investigation using the dataset provided by the GHTorrent
project~\cite{Gousi13}, an off-line mirror of the data offered through the
Github API. The Github API data come in two forms; a streaming data flow lists
events, such as reporting an issue, happening on repositories in real time,
while a static view contains the current state of entities. The data is stored
in unprocessed format, in a MongoDB database, while metadata is extracted and
stored in a relational database. The GHTorrent dataset covers a broad range of
development activities on Github, including issues. The project has been
collecting data since February 2012. Up to Dec 2013, when we queried it to
retrieve our bootstrapping sample, it included 356.057 issues from 16.836
Java projects (step 1).\footnote{A Java project is specified in GitHub with a
specific label.}

@@

\noindent\emph{Kinds of Analyzed Issues.} Issues on Github are different from
issues on dedicated bug tracking tools such as Bugzilla and Jira. The most
important difference is that there are no predefined fields for bug reporters to
select from (e.g. severity and priority); Github uses instead a more open ended
tagging system. Repositories are offered a pre-defined set of labels, but
repository owners can modify them at will. An issue can have an arbitrary set of
labels attached. Hence, this study does not restrict the analysis of stack
traces defined on defect-issues, based on the assumption that: regardless the
type of issue where its is defined the stack trace information contain relevant
information concerning the exception use. Moreover, issue and pull requests are
dual on Github; all pull requests have a corresponding ``backing'' issue which
are automatically generated. Therefore, we needed to exclude pull request based
issues from our analysis. Finally, the Github API allows the automated
generation of issues for specific repositories, which automated tools often use
to report crashes. In some cases, this led to high number of issues that
included stack traces. We identified and filtered those cases out (e.g.,
the \textsf{pullwifi} project was responsible for almost 50\% of all issues created on
GitHub, it was removed from our analysis and also from GitHub).

@@

\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
is classifying exceptions as either checked exceptions, runtime exceptions or an
errors. The exception classifier is based on bytecode analysis (based on Design
Wizard~\cite{Brunet09}) that walks up the type hierarchy of a Java exception
until it reaches a base exception type.

@@

The use of the Github issue tracker in repositories hosted on Github is optional
and not widespread. As of March 2014, from the 332,864 non-fork Java
repositories registered in the GHTorrent dataset, only 44,323 have received an
issue report in their lifetime while from those only 16,837 are being watched by
external users. To ensure that our work targets projects that are openly used by
users other than their developers, we focused our analysis on just those
projects. Using the heuristics presented in Section~\ref{sec:android}, we
selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
Table~\ref{tab:overview} presents two views of the extracted dataset, namely
grouped by repository and grouped by issue.

From Table~\ref{tab:overview}, we can see that repositories that include at
least one exception or stack trace on an issue are fairly common: 28\% and 22\%
respectively. Considering only the exceptions and stack traces reported on
defect issues we find out that 41\% of the repositories reported at least one
exception in a defect issue, and approximately 35\% of the projects mentioned
one stack trace in at least one defect issue.

@@

Programming mistakes are the most common root causes for most types of signalers. We
could observe that the NullPointerException is still the exception with higher
number of occurrences. The NullPoiterExceptions are mainly signaled inside
Android platform and inside application code, although we also find
NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus if it is a good or bad practice to re-throw
NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

We could observe in this study that the NullPointerException and other
other implicitly signaled exceptions represented most of the exceptions reported
on issues in both Java and Android subset. For such exceptions, which represent
programming bugs or resource limitations, there is usually no proper handling
besides presenting an error message to the user and restarting the application ---
only high fault tolerant systems need to provide solutions to handle them.

@@

Throwing general exceptions, such as direct instances of RuntimeException, is  considered a
bad practice, because the exception type does not carry enough information to identify the
cause of the exceptional behaviour, and as a consequence developers need to relying on
the exception message which can be neither complete nor precise
~\cite{gosling2000java}.

@@

To better investigate the extent of this problem we classified each stack trace according to the type of the
root exception. Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on 
the Android repositories. As we can see from Table ~ref{typeroottab},  most of the exceptions signaled by third-party code
 (e.g., Android and libs) were runtime exceptions. After filtering all the exceptions implicitly signaled by 
JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an inderect 
instances of java.lang.RuntimeException). We could observe observe that only 1 method out of 118 inspected signalers
 (i.e., 0,8%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other  study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

Such undocumented runtime exceptions represent a serious threat to system robustness, specially
when such exceptions are thrown then the third party code is invoked inside the application (e.g. libraries, or framework utility code).
In such cases the client usualy do not have access to the source code, and in the absense of
the exception documentation it is very difficult or even impossible for the client of such thrid party code to 
protect the application against undocumented runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

@@

On the other hand, we could observe that most of the exceptions signaled by the set of 
libraries reused by Android platform (i.e., org.apache.harmony,
org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax -- referred in
this work as libcore) were checked exceptions. It is pointed as a good practice
for libraries (see Section~\ref{sec:best}) because by using checked exceptions
libraries can define a precise exception interface to its clients. Such
libraries are heavily used in several projects, and such precise exception
interface may be related to the libraries maturity.

@@

Although most of the  reported exceptions were runtime, we could also find checked
exceptions as the root causes of stack traces reported on issues. A checked exception
can remain uncaught in two circumstances: if all methods on the execution trace
on which this exception flows explicitly specifies the exception, or if the
checked exception is wrapped in a runtime and than re-thrown.

Since checked exceptions should be used to represent recoverable conditions, ignoring a
checked exception, is considered a bad practice. As mentioned before, when the developer is confronted
with a checked exception, the designer of the API is telling him to handle the
exceptional condition. To better understand what was causing checked exceptions
to scape from being handled we investigated the kinds of wrappings that happened
on stacks the stacks.  Next sections presents the wrappings found and discuss
about them.

@@

This analysis also revealed unexpected wrappings such as: Error wrapping a Checked 
exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
Java is the only language that provides a hybrid exception handling mechanism
which offers  different types to represent different exception behaviors (i.e., error, 
runtime and checked) (see Section ~\ref{sec:best}) . According to Java specification Errors should not be
handled inside the system since they usually represent unrecoverable conditions
detected by the JVM such as OutOfMemoryError. Checked exceptions should
represent recoverable conditions and  Runtime exceptions represent are usually
related to programming mistakes from which the developer is not expected to recover
from.

@@

\section{Discussions}

The exception (mis-)use patterns that emerged  in this study sheds light on 
problems that may be reducing the robustness of current Java programs in the face of
exceptional conditions. \noindent\emph{The undocumented runtime exceptions} thrown by library code -
documenting runtime exceptions is a tedious and error prone task, to help developers
mitigating this problem tools should be developed to automate document runtime exceptions
scaping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}. 
 \noindent\emph{The odd wrappings}  found in this study can lead to what we call
an \emph{exception handling confusion problem} which can lead the program to
an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
 one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the JVM which make impossible the program to recover. Further investigation is needed on finding 
ways to help developers in deadling with such problem, either preventing odd wrappings or enabling 
the developer to better dealing with them. Last but not least, the high number of null pointer exceptions 
lead us to think on other pertinent research direction: whether language designs which avoid null pointers, 
such as Monads ~cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

@@

In this section, we present works that are related to our own, distributed in
four categories: i) works that use the information available on stack traces;
ii) empirical studies on the usage of Java exceptions and its fault proneness;
and iii) tools to extract stack traces information from natural language artifacts
(i.e., issues and emails).
%, and iv) empirical studies involving Android apps.

\textit{Analysis of Java Exception Usage.} Since the manual analysis of the Java
exception flow can easily become infeasible, Robillard and Murphy~\cite{Robil00}
employed dataflow analysis to find the propagation paths of checked and
unchecked exception types. Modeled after Robillard and Murphy's work, other
tools have been proposed to support the static analysis of exception
flows~\cite{coelho2008assessing}. The main limitation off all
static analysis tools is the number of false positives inherent to static analysis
solutions, which can lead to a high number of exception flows, specially if
considering Java Environment exceptions and exceptions signaled from libraries.
Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
exception handling code of 32 open-source systems, both for Java and .NET. They
observed that the action handlers were very simple (e.g., logging and present a
message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
set of bad practices on exception handling that hinder software maintainability
and robustness, based on their own experience with Java enterprise applications.
Our work differs from those two works as we tried to identify bad practices from
the combined use of stack traces extracted from issues and bytecode and source
code analysis.

@@

\textit{Extracting Stack Traces from natural language artifacts.} 
%Currently, many
%software vendors embed automatic crash reporting tools in their software
%systems. Moreover, third party crash collection services exist, most of them 
%targeted for applications run on mobile phones~\cite{BugSe14,BugSn14,Googl14,Acra14}.
Apart from bug reports, stack traces can be embedded in other forms of
communication between developers, such as discussion logs and emails.
Being intermixed with text makes the accurate extraction of stacktraces 
an involved process.
Infozilla~\cite{bettenburg2008extracting}
is based on a set of regular expressions that extract a set of frames
related to a stack trace. The main limitation of this solution is that it is not
able to extract stack traces embedded on verbose log files (i.e., on which we
can find log text mixed with exception frames). Bacchelli and Lanza
~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
from development emails and relate it to code artifacts (i.e. classes) mentioned
on the stack trace. In addition to those tools, ExceptionMiner is able to 
both extract stack traces from natural language artifacts and to 
classify them in a set of predefined categories.

>>>>

Modern applications have to cope with an increasing number of abnormal
computation states that arise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware. The exception handling
mechanism~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions. Although the
exception handling mechanism has been embedded in several programming languages
(e.g. Java, C++, C\#) and was the target of several studies
(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
the exception handling code is often generally poorly understood and the least
tested part of software systems.

Often exception handling constructions may lead the developers into believing that
by just re-throwing the exceptions they can forget about the exceptional
situations during the development of the "happy path". This "ignore-for-now"
approach may turn the exception handling into a generalized "goto"
mechanism~\cite{mandrioli1992advances} making the program more complex and even
less reliable. This behaviour may lead to the well known uncaught exceptions
problem~\cite{jo2004uncaught}, one of the main causes of application crashes.

Moreover, some languages use exceptions to represent programming mistakes
(out-of-bounds array index, division-by-zero, access to null pointers). In these cases,
exceptions are implicitly signaled by the runtime environment when such
conditions occur. One of such implicit exceptions (caused by data conversion
from a 64-bit floating point to a 16-bit signed integer) was the responsible for
the famous failure on Ariane 5's first test flight~\cite{lions1996ariane} ---
leading to the rocket self-destructing 37 seconds after launch and a loss of 500
million dollars. Besides Ariane 5, several applications crash everyday due to
uncaught exceptions~\cite{jo2004uncaught}. Not using exceptions carefully
may have the opposite effect of the initial intention of improving
system robustness.

@@

To answer these questions, we conduct an exploratory study of stack traces in
the wild. Using a custom tool called ExceptionMiner, that was specifically
developed to for this study, we mined stack traces from issues across all Java
projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
from which 28,800 stack traces were extracted and used as a source of
information for our analysis. The stack trace analysis  was augmented by
additional information extracted using byte code and source code analysis done
on a carefully selected subset of our initial sample (482 Android projects).
%This study was divided in two phases: in the fist phase all the issues of all
%Java projects available on GitHub were analyzed and the embedded stack traces
%extracted using ExceptionMiner.

@@

  \item  It introduces ExceptionMiner, a tool developed to support the
    analysis.

@@

The contributions of this work allow for developers of Java applications
to familiarize with the most common reported exceptions mis-uses which is the
first step to help developers to avoid making them; (ii) designers of languages
to consider ways of reducing the abundance of NullPointerExceptions, (iii) and
tool designers to consider developing tools that enable a cross-project defect
analysis. We strongly believe that, specially in the open-source environment,
faults are not to be hidden in a private bug issue. Faults should be shared and
discussed, so that a developer can learn from other projects mistakes.
Currently, the search facilities of repositories are very limited, and the
ExceptionMiner tool is a contribution in this direction.

@@

The Java specification~\cite{gosling2000java} suggests that user-defined
exceptions should be checked exceptions. The main reason is because doing so the
callers of a method will know about the exceptions that a method can throw, and
so they can decide what to do about them. However, there is a long-lasting
debate about the use of checked and unchecked exceptions in
Java~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
are related to this debate} since there are pros and cons associated to each of
them.

@@

Several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. 
A list of the guidelines related to exception types and exception
propagation in Java the following:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

@@

\textbf{I-Use checked exceptions to represent recoverable
conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
The developer should use checked exceptions for conditions from which the caller
is expected to recover. By confronting the API user with a checked exception,
the API designer is forcing the client to handle the exceptional condition. The
client can explicitly ignore the exception (swallowing, or converting it to
other type) at the expense of the program's robustness~\cite{gosling2000java}.

@@

%\noindent\emph{Exception Documentation}
\textbf{IV-Document all exceptions thrown by a
method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled JVM runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of all the exceptions the method can throw. If the developer fails to
do follow this practice (specially when developing library code) it will be
difficult or even impossible for the caller to make effective use of such 
method~\cite{wirfs2006toward, bloch2008effective}.

@@

This work describes an exploratory study which was based on a sequential
mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and
integrating both quantitative and qualitative evidence along the stages of the
research process. The main goal of this study was to gain a better understanding
of the use of exceptions in Java programs based mainly on the information
available on exception stack traces. More specifically, we aimed at answering
the following research questions:

@@

For RQ1 and RQ2, we explore the domain quantitatively, and use theirs answers as
a basis to investigate the third research question. RQ3 aims at examining whether
exception (mis)-uses could be revealed through stack trace analysis. To
support this investigation, source code and bytecode analysis was used to
leverage the  understanding of stack traces. 
%For RQ3 we investigate
%quantitatively and highlight interesting cases by exploring cases qualitatively.
Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
conducted to answer these questions.

@@

We begun our investigation using the dataset provided by the GHTorrent
project~\cite{Gousi13}, an off-line mirror of the data offered through the
Github API. The Github API data come in two flavours; a streaming data flow lists
events, such as reporting an issue, happening on repositories in real time,
while a static view contains the current state of entities. The data is stored
in unprocessed format, in a MongoDB database, while metadata is extracted and
stored in a relational database. The GHTorrent dataset covers a broad range of
development activities on Github, including issues. The project has been
collecting data since February 2012. Up to Dec 2013, when we queried it to
retrieve our bootstrapping sample, it included 356.057 issues from 16.836
Java projects (step 1).\footnote{A Java project is specified in GitHub with a
specific label.}

@@

\noindent\emph{Kinds of Analyzed Issues.} Issues on Github are different from
issues on dedicated bug tracking tools such as Bugzilla and Jira. The most
important difference is that there are no predefined fields for bug reporters to
select from (e.g. severity and priority); Github uses instead a more open ended
tagging system. Repositories are offered a pre-defined set of labels, but
repository owners can modify them at will. An issue can have an arbitrary set of
labels attached. Hence, this study does not restrict the analysis of stack
traces defined on defect-issues, based on the assumption that regardless of the
type of issue where its is defined the stack trace information contain relevant
information concerning the exception use. Moreover, issues and pull requests are
dual on Github; all pull requests have a corresponding ``backing'' issue which
are automatically generated. Therefore, we excluded pull request based
issues from our analysis. Finally, the Github API allows the automated
generation of issues for specific repositories, which automated tools often use
to report crashes. In some cases, this led to high number of issues that
included stack traces. We identified and filtered those cases out (e.g.,
the~\textsf{pullwifi} project was responsible for almost 50\% of all Android issues in our dataset).

@@

\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
is classifying exceptions as either checked exceptions, runtime exceptions or 
errors. The exception classifier is based on bytecode analysis (based on Design
Wizard~\cite{Brunet09}) that walks up the type hierarchy of a Java exception
until it reaches a base exception type.

@@

%The use of the Github issue tracker in repositories hosted on Github is optional
%and not widespread. As of March 2014, from the 332,864 non-fork Java
%repositories registered in the GHTorrent dataset, only 44,323 have received an
%issue report in their lifetime while from those only 16,837 are being watched by
%external users. To ensure that our work targets projects that are openly used by
%users other than their developers, we focused our analysis on just those
%projects. Using the heuristics presented in Section~\ref{sec:android}, we
%selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
%Table~\ref{tab:overview} presents two views of the extracted dataset, namely
%grouped by repository and grouped by issue.
%
From Table~\ref{tab:overview}, we can see that repositories that include at
least one exception or stack trace on an issue are fairly common: 28\% and 22\%
respectively. Considering only the exceptions and stack traces reported on
defect issues we find out that 41\% of the repositories reported at least one
exception in a defect issue, and approximately 35\% of the projects mentioned
one stack trace in at least one defect issue.

@@

Programming mistakes are the most common root causes for most types of signalers. We
could observe that the NullPointerException is still the exception with higher
number of occurrences. The NullPoiterExceptions are mainly signaled inside
Android platform and inside application code, although we also find
NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw
a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

We could observe in this study that the NullPointerException and other
other implicitly signaled exceptions represented most of the exceptions reported
on issues in both Java and Android subset. For such exceptions, which represent
programming bugs or resource limitations, there is usually no proper handling
besides presenting an error message to the user and restarting the application.
Only high fault tolerant systems need to provide solutions to handle them.

@@

Throwing general exceptions, such as direct instances of RuntimeException, is  considered a
bad practice, because the exception type does not carry enough information to identify the
cause of the exceptional behaviour, and as a consequence developers need to relying on
the exception message which can be neither complete nor precise~\cite{gosling2000java}.

@@

To better investigate the extent of this problem we classified each stack trace according to the type of the
root exception. Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on 
the Android repositories. As we can see from Table ~ref{typeroottab},  most of the exceptions signaled by third-party code
 (e.g., Android and libs) were runtime exceptions. After filtering all the exceptions implicitly signaled by 
JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an inderect 
instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other  study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

Such undocumented runtime exceptions represent a threat to system robustness, specially
when such exceptions are thrown then the third party code is invoked inside the application (e.g. libraries, or framework utility code).
In such cases the client usually do not have access to the source code, and in the absence of
the exception documentation it is very difficult or even impossible for the client of such third party code to 
protect the application against undocumented runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

@@

On the other hand, we could observe that most of the exceptions signaled by the set of 
libraries reused by Android platform (i.e., org.apache.harmony,
org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax -- referred in
this work as libcore) were checked exceptions. It is pointed as a good practice
for libraries (see Section~\ref{sec:best}) because by using checked exceptions
libraries can define a precise exception interface to its clients. Such
libraries are heavily used in several projects, and such precise exception
interface may be related to the library's maturity.

@@

Although most of the  reported exceptions were runtime, we could also find checked
exceptions as the root causes of stack traces reported on issues. A checked exception
can remain uncaught in two circumstances: if all methods on the execution trace
on which this exception flows explicitly specifies the exception, or if the
checked exception is wrapped in a runtime and than re-thrown. Since checked exceptions 
should be used to represent recoverable conditions, ignoring a
checked exception, is considered a bad practice. As mentioned before, when the developer is confronted
with a checked exception, the designer of the API is telling him to handle the
exceptional condition. To better understand what was causing checked exceptions
to scape from being handled we investigated the kinds of wrappings that happened
on stacks the stacks.  Next sections presents the wrappings found and discuss
about them.

@@

This analysis also revealed unexpected wrappings such as: Error wrapping a Checked 
exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
Java is the only language that provides a hybrid exception handling mechanism
which offers  different types to represent different exception behaviors (i.e., error, 
runtime and checked) (see Section~\ref{sec:best}). According to Java specification Errors should not be
handled inside the system since they usually represent unrecoverable conditions
detected by the JVM such as OutOfMemoryError. Checked exceptions should
represent recoverable conditions and  Runtime exceptions represent are usually
related to programming mistakes from which the developer is not expected to recover
from.

@@

\section{Discussion}

The exception (mis-)use patterns that emerged  in this study sheds light on 
problems that may be reducing the robustness of current Java programs in the face of
exceptional conditions. \noindent\emph{The undocumented runtime exceptions} thrown by library code -
documenting runtime exceptions is a tedious and error prone task, to help developers
mitigate this problem tools should be developed to automate the documentation of runtime exceptions
scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}. 
 \noindent\emph{The odd wrappings}  found in this study can lead to what we call
an \emph{exception handling confusion problem} which can lead the program to
an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
 one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the JVM which make impossible the program to recover. Further investigation is needed on finding 
ways to help developers in deadling with such problem, either preventing odd wrappings or enabling 
the developer to better dealing with them. Last but not least, the high number of null pointer exceptions 
lead us to think on other pertinent research direction: whether language designs which avoid null pointers, 
such as Monads ~cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

@@

In this section, we present works that are related to our own, distributed in
three categories: i) works that use the information available on stack traces;
ii) empirical studies on the usage of Java exceptions and its fault proneness;
and iii) tools to extract stack traces information from natural language artifacts
(i.e., issues and emails).
%, and iv) empirical studies involving Android apps.

\textit{Analysis of Java Exception Usage.} As the manual analysis of the Java
exception flow can easily become infeasible, Robillard and Murphy~\cite{Robil00}
employed dataflow analysis to find the propagation paths of checked and
unchecked exception types. Modeled after Robillard and Murphy's work, other
tools have been proposed to support the static analysis of exception
flows~\cite{coelho2008assessing}. The main limitation of all
static analysis tools is the number of false positives inherent to static analysis
solutions, which can lead to a high number of exception flows, specially if
considering Java Environment exceptions and exceptions signaled from libraries.
Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
exception handling code of 32 open-source systems, both for Java and .NET. They
observed that the action handlers were very simple (e.g., logging and present a
message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
set of bad practices on exception handling that hinder software maintainability
and robustness, based on their own experience with Java enterprise applications.
Our work differs from those two works as we tried to identify bad practices from
the combined use of stack traces extracted from issues and bytecode and source
code analysis.

@@

\textit{Extracting Stack Traces from natural language artifacts.} 
%Currently, many
%software vendors embed automatic crash reporting tools in their software
%systems. Moreover, third party crash collection services exist, most of them 
%targeted for applications run on mobile phones~\cite{BugSe14,BugSn14,Googl14,Acra14}.
Apart from bug reports, stack traces can be embedded in other forms of
communication between developers, such as discussion logs and emails.
Being intermixed with text makes the accurate extraction of stacktraces 
an involved process.
Infozilla~\cite{bettenburg2008extracting}
is based on a set of regular expressions that extract a set of frames
related to a stack trace. The main limitation of this solution is that it is not
able to extract stack traces embedded on verbose log files (i.e., on which we
can find log text mixed with exception frames). Bacchelli
et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
from development emails and relate it to code artifacts (i.e. classes) mentioned
on the stack trace. In addition to those tools, ExceptionMiner is able to 
both extract stack traces from natural language artifacts and to 
classify them in a set of predefined categories.
